### 2020-04-15 ICT_Frontiers_rev202001.tex
Updates from Overleaf

% Three major changes
% 1. 2.1
% 2. 4.1  non-monotonic 
% 3. 5.1 Processes that are not Informationally Closed
% 4. appendix

% todo
% ------------------------------------
% unpredictable sensory 
% As mentioned above, informational closure can happen between scales of coarse-graining within a single system.....

\documentclass[utf8]{article}
\usepackage{import}

%\ProvidesPackage{WritingToolsByAcer}

%% Language and font encoding
\usepackage[english]{babel}
%\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}

%% Sets page size and margins
\usepackage[a4paper, top=3cm,bottom=2cm,left=3cm,right=3cm,marginparwidth=4cm]{geometry}
% \usepackage[papersize={10in, 12in}, top=3cm,bottom=2cm,left=2in,right=2in,marginparwidth=1.8in]{geometry}

%% Title style
\usepackage{stringstrings}
\usepackage[explicit]{titlesec}
%\usepackage{sectsty}

%% ============================== Useful packages ============================= %
\usepackage[dvipsnames,table,xcdraw]{xcolor}
\usepackage[round, sort]{natbib}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{graphicx}
\usepackage{svg}
\usepackage[colorlinks=true, allcolors=blue]{hyperref}
\usepackage{authblk}
\usepackage{float}
\usepackage{tikz}
\usepackage{lineno}
\usepackage{csquotes}

% =================================== Macro ================================== %
\usepackage[colorinlistoftodos]{todonotes}
\usepackage[at]{easylist}
\usepackage{tcolorbox}
\tcbuselibrary{skins}
\usepackage{comment}
\usepackage[framemethod=tikz]{mdframed}
\usetikzlibrary{shadows}
\usepackage{soul}
\usepackage{version}

% Meaning of colours --------------------------------------------------------- %
%	Red: Important and Critical
%	Orange: Todo
%	Blue: Topic and Guideline
%	Purple: Murmur
%	Green: Question, Call for help, Open for opinions and comments
%	Brown: Need Wording
% ---------------------------------------------------------------------------- %

% Real content
	% Outline items that can be integrated into main text
		\newenvironment{ants}
			{
			 \begin{easylist}[itemize]
		 	}
			{
			\end{easylist}
			} %

	% Writing Materials
    \newenvironment{WritingMaterials} %
    	{
            \begin{tcolorbox}[enhanced,
                title=-,
                size=small,
                colbacktitle=Aquamarine,
                drop fuzzy shadow,
                fontupper=\small,
                boxrule=0.4pt,
                colback=Aquamarine!10!white,
                sharp corners]
                Writing Materials
            \end{tcolorbox}
            \begin{easylist}[itemize]
    	}
    	{
            \end{easylist}  
            \begin{tcolorbox}[enhanced,
                halign=flush right,
                halign title=right,
                size=small,
                colbacktitle=Aquamarine,
                drop fuzzy shadow,
                fontupper=\small,
                boxrule=0.4pt,
                colback=Aquamarine,
                colupper=White,
                sharp corners]
                Writing Materials -
            \end{tcolorbox}        
    	}

    % \excludeversion{WritingMaterials}

% Functional content
	% Quotes: green
		\newcommand{\rewrite}[1]{\textcolor{ForestGreen}{\textit{"#1"}}\newline}

	% todo: orange
		% text todo
		\newcommand{\toWrite}[1]{\noindent
			\textcolor{Orange}{\textbf{[[ #1 ]]}}}

	% figure todo
        \newcommand{\includegraphicsTodo}[2][]{%
            \tcbox[enhanced,
                center,
            	adjusted title=\large To Be Modified \dots,
            	halign title=center,
            	attach boxed title to top left={yshift=-3mm, xshift=1cm,yshifttext=-1mm}, boxed title style={drop fuzzy shadow},
            	colbacktitle=RedOrange!70!black,
            	coltitle=white,
            	colframe=RedOrange!50!black,
            	colback=white
        	]{\includegraphics[#1]{#2}}}

		% figure todo callout
        \newcommand{\needfig}[1]{%
			\ifthenelse{\equal{#1}{}}{%
				\todo[color=White, linecolor=Orange, bordercolor=RedOrange]{\textcolor{RedOrange}{Fig}}}{%
				\todo[color=White, linecolor=Orange, bordercolor=RedOrange]{\textcolor{RedOrange}{Fig: #1}}%
			}%
		}

		% rewording
		\newcommand{\rewording}[1]{\textcolor{RawSienna}{[[ REWORDING | #1 ]]}}

		% Need reference
		\newcommand{\needref}[1]{%
			\ifthenelse{\equal{#1}{}}{%
				\todo[color=White, linecolor=Orange, bordercolor=Orange]{\textcolor{Orange}{Ref}}}{%
				\todo[color=White, linecolor=Orange, bordercolor=Orange]{\textcolor{Orange}{Ref: #1}}%
			}%
		}

	% Highlight: Yellow or Red
		% Highlight only
		\newcommand{\highlight}[2][Yellow]{\sethlcolor{#1}\hl{#2}}

		% Highlight and comment
        \newcommand{\highlightAndComment}[3][Yellow]{\sethlcolor{#1}\hl{#2}\todo[color=#1, linecolor=Yellow]{#3}}

		% Critical point
		\newcommand{\critical}[1]{%
			\todo[color=OrangeRed!80!white]{#1}
		}

	% idea: Purple
		% text idea
			\newcommand{\idea}[2][Plum]{\noindent
				\textcolor{#1}{[[ #2 ]]}}

		% box idea
			\newcommand{\ideaBox}[1]{
			    \vspace{2mm}
				\begin{tcolorbox}[enhanced,%
                    title=-,
                    center,
                    fontupper=\small,
                    lifted shadow={1mm}{-2mm}{3mm}{0.1mm}{black!50!white},%
                    boxrule=0.4pt,
                    width=9cm,
                    colback=Thistle!50!white,
                    sharp corners]
					#1
				\end{tcolorbox}
				\vspace{2mm}
			}

		% call-out idea
			\newcommand{\ideaCallout}[1]{\todo[color=Thistle!50!white, size=\small]{#1}}

	% communication: green
		% call for help
			\newcommand{\needhelp}[1]{\todo[color=SpringGreen]{[Need Help]: #1}}

		% comment
			% comment by authors
			\newcommand{\addcomment}[3][\empty]{\todo[color=SkyBlue]{Comment [#2]:\newline#3}\colorbox{SkyBlue}{#1}}

			% usage:
			% \addcomment{auther}{comment text}
			% or
			% \addcomment[highlighted text]{auther}{comment text}
			%
			% e.g.
			% It's a \addcomment[beautiful]{Acer}{need a reference} day

    % backup: 
    \newenvironment{backup}
    {

        \begin{tcolorbox}[enhanced,
            title=-,
            size=small,
            colbacktitle=black!70!white,
            colback=black!10!white,            
            drop fuzzy shadow,
            fontupper=\small,
            boxrule=0.4pt,
            sharp corners]
            Backup
        \end{tcolorbox}
    }
    {
        \begin{tcolorbox}[enhanced,
            halign=flush right,
            halign title=right,
            size=small,
            colbacktitle=black!10!white,
            colback=black!70!white,
            drop fuzzy shadow,
            fontupper=\small,
            boxrule=0.4pt,
            colupper=White,
            sharp corners]
            Backup -
        \end{tcolorbox}    
    }
    % \excludeversion{backup}

% review loop annotation
    \usepackage[xcolor, outerbars]{changebar}
    \cbcolor{red}

    % start anchor: \rlstart
    \newcommand{\rlstart}[1]{  %
        \setlength\changebarwidth{#1pt*2}  %
        \hspace*{-40pt}
        \cbstart{\textcolor{red}{\newline\textbf{iteration #1}}\newline}}

    % end anchor: \rlend
    \newcommand{\rlend}{\cbend}

% quote
\newcommand{\myQuote}[1]{\textit{``#1''}}

% Review ---------------------------------------------------------------------------- %
\newenvironment{reply}  
    {\color{Blue}\noindent\newline}
    {\newline}

% \revise
%   {Page}{original text}{revised text}
\newcommand{\revise}[2]{
    \noindent
    \newline
    Original:\newline
    \textit{"#1"}
    \newline
    \newline
    Revised:\newline
    \textit{"#2"}\newline}

% \excludeversion{WritingMaterials}
% \excludeversion{backup}

\usepackage{amsthm}

% list of fig para
\usepackage{tocloft}
\newlength{\mylen}
\renewcommand{\cftfigpresnum}{\figurename\enspace}
\renewcommand{\cftfigaftersnum}{:}
\settowidth{\mylen}{\cftfigpresnum\cftfigaftersnum}
\addtolength{\cftfignumwidth}{\mylen}

% \theoremstyle{definition}
\newtheorem{definition}{Definition}
\newtheorem*{hypothesis}{Hypothesis}

\usepackage{chngcntr}
\counterwithout{figure}{section}

%% ================================== acronym ================================= %
\usepackage{acronym}
\newacro{ntic}[NTIC]{non-trivial informational closure}
\newacro{ic}{informational closure}
\newacro{cg}{coarse-grain}
\newacro{cging}{coarse-graining}
\newacro{cged}{coarse-grained}
\newacro{ncg}{neural coarse-graining}
\newacro{OurTheory}[ICT]{Information Closure Theory of Consciousness}

\newcommand{\X}{\mathcal{X}}
\renewcommand{\S}{\mathcal{S}}
\newcommand{\E}{\mathcal{E}}
\DeclareMathOperator*{\argmax}{arg\,max}
% ============================================================================ %
%                                     Title                                    %
% ============================================================================ %
\title{Information Closure Theory of Consciousness}
\date{}

% ============================================================================ %
%                                    Authors                                   %
% ============================================================================ %
\author[]{Acer Y.C. Chang\thanks{Corresponding author, acercyc@araya.org}}
\author[]{Martin Biehl\thanks{martin@araya.org}}
\author[]{Yen Yu\thanks{yen.yu@araya.org}}
\author[]{Ryota Kanai\thanks{kanair@araya.org }}
\affil[]{ARAYA, Inc., Tokyo, Japan}

% ============================================================================ %
%                                     Body                                     %
% ============================================================================ %
\begin{document}
    \linenumbers
	\maketitle
	\tableofcontents

	\begin{abstract}
		Information processing in neural systems can be described and analysed at multiple spatiotemporal scales. Generally, information at lower levels is more fine-grained and can be coarse-grained in higher levels. However, information processed only at specific scales of coarse-graining seems to be available for conscious awareness. We do not have direct experience of information available at the scale of individual neurons, which is noisy and highly stochastic. Neither do we have experience of more macro-scale interactions such as interpersonal communications. Neurophysiological evidence suggests that conscious experiences co-vary with information encoded in coarse-grained neural states such as the firing pattern of a population of neurons. In this article, we introduce a new informational theory of consciousness: \acf{OurTheory}. We hypothesise that conscious processes are processes which form \ac{ntic} with respect to the environment at certain coarse-grained scales. This hypothesis implies that conscious experience is confined due to informational closure from conscious processing to other coarse-grained scales. \ac{OurTheory} proposes new quantitative definitions of both conscious content and conscious level. With the parsimonious definitions and a hypothesise, \ac{OurTheory} provides explanations and predictions of various phenomena associated with consciousness. The implications of \ac{OurTheory} naturally reconciles issues in many existing theories of consciousness and provides explanations for many of our intuitions about consciousness. Most importantly, \ac{OurTheory} demonstrates that information can be the common language between consciousness and physical reality.

	\end{abstract}

	\section*{Keywords:}
	Keywords: theory of consciousness, non-trivial informational closure, NTIC, coarse-graining, level of analysis

% ============================================================================ %
%                                     Start                                    %
% ============================================================================ %

    \newpage
	\section{Introduction}

		%% Story about a cell
		Imagine you are a neuron in Alice's brain. Your daily work is to collect neurotransmitters through dendrites from other neurons, accumulate membrane potential, and finally send signals to other neurons through action potentials along axons. However, you have no idea that you are one of the neurons in Alice's supplementary motor area and involved in many motor control processes for Alice's actions, for example, grabbing a cup. You are ignorant of intentions, goals, and motor plans that Alice has at every moment even though you are part of the physiological substrate responsible for all those actions.
		A similar story also happens to Alice's conscious mind. To grab a cup, for example, Alice is conscious of her intention and visuosensory experience of this action. However, her conscious experience does not reflect the dynamic of your membrane potential or the action potentials you send to other neurons every second. That is, not all information you have is available to Alice's conscious mind.

		%% scale

		It seems to be true that we are not consciously accessing information processed at every scale in the neural system. There are both more microscopic and more macroscopic scales than the scale corresponding to the conscious contents. On the one hand, dynamics of individual neurons are stochastic \citep{Goldwyn2011, White2000}. However, what we are aware of in our conscious mind shows astonishing stability and robustness against the ubiquitous noise in the neural system \citep{mathis1995computational}. In addition, some parts of the neural system contribute very little to conscious experience (the cerebellum for example \citep{lemon2010life}), also suggesting that conscious contents do not have one-to-one mapping to the entire state of the neural system. On the other hand, human conscious experience is more detailed than just a simple (e.g. binary) process can represent, suggesting that the state space of conscious experience is much larger than what a single overly coarse-grained binary variable can represent. These facts suggest that conscious processes occur at a particular scale. \linelabel{line:lack-theory}We currently have only few theories (e.g., Integrated Information Theory \citep{hoel2016can} and Geometric Theory of Consciousness \citep{fekete2011towards,fekete2012lack}) to identify the scale which conscious processes correspond to ( also see discussion in \cite{fekete2016system}). We refer to this notion as \textbf{the scale problem of consciousness} (Fig.~\ref{fig:scaleproblem}).

		%% our argument
		In this article, we propose a new information-based theory of consciousness, called \acf{OurTheory}. We argue that every process with a positive non-trivial information closure (NTIC) has consciousness. This means that the state of such a process corresponds one-to-one to conscious content.\footnote{In the following IC stands for "informational closure" or "informationally closed" and NTIC stands for "non-trivial informational closure" or "non-trivially informationally closed".}. We further postulate that the \textit{level} of consciousness corresponds to the degree of NTIC. (for a discussion of the distinction between level versus content of consciousness see \cite{laureys2005neural, overgaard2010neural}).

		In the following, we first introduce non-trivial informational closure and argue for its importance to information processing for human scale agents (Sec.\ref{sec:Non-trivial informational closure}). We next argue that through coarse-graining the neural system can form a high degree of NTIC at a specific coarse-grained scale (Sec.\ref{sec:Neural coarse-graining}). In the Sec.\ref{sec:OurTheory}, we propose a new theory of consciousness (\ac{OurTheory}). We also illustrate how \ac{OurTheory} can parsimoniously explain empirical findings from previous consciousness studies (Sec.\ref{sec:Conscious versus Unconscious Processing}) and reconcile several current major theories of consciousness (Sec.\ref{sec:Comparison with other theories}). Finally, we discuss the current theoretical and empirical limitations of \ac{OurTheory} and propose the implications from \ac{OurTheory} to the current consciousness science (Sec.\ref{sec:Limitation and Future work}).

		\begin{figure}[H]
		    \centering
			\includegraphics[width=\textwidth]{WritingMaterials/Fig_ScaleProblemOfConsciousness/ScaleProblemOfConsciousness.pdf}
			\caption{The scale problem of consciousness: Human conscious experience does not reflect information from every scale. Only information at a certain coarse-grained scale in the neural system is reflected in consciousness.}
			\label{fig:scaleproblem}
	   	\end{figure}

% ============================================================================ %
%                       Non-trivial informational closure                      %
% ============================================================================ %
	\section{Non-trivial Informational Closure} \label{sec:Non-trivial informational closure}
		The notion of non-trivial informational closure (NTIC) is introduced by \cite{BERTSCHINGER.2006}. The concept of closure is closely related to system identification in systems theory. One can distinguish a system from its environment by computing the closedness of the system \citep{maturana1991autopoiesis, rosen1991life, pattee2012evolving, luhmann1995probleme}. The closedness can be further quantified by information theory.

		% Definition of informational closure
			\linelabel{line:r1q1a}Consider two processes, the environment process $(E_t)_{t \in \mathbb{N}}$ and the system's process $(Y_t)_{t \in \mathbb{N}}$ and let their interaction be described by the Bayesian network in Fig.~\ref{fig:SystemAndEnv}. Then, information flow $J_{t}$ from the environment $E$ to a system $S$ at time $t$ can be defined as the conditional mutual information $I$ between the current environment state $E_{t}$  and the future system state $Y_{t+1}$ given the current system state $Y_{t}$

				\begin{equation}
    				\label{eq:InformationFlow}
    				\left.\begin{array}
    				{rl}{J_{t}(E \rightarrow Y )} & {:= I(Y_{t+1};E_{t}|Y_{t})} \\
    				{ } & { \ = I(Y_{t+1};E_{t}) - (I(Y_{t+1};Y_{t})-I(Y_{t+1};Y_{t}|E_{t}))}
    				\end{array}\right.
				\end{equation}

				\begin{figure}
					\includegraphics[width=\textwidth]{Submission/Frontiers/SystemAndEnv_3.pdf}
					\caption{The dependencies between a system  $Y$ and its environment $E$ through the channels $\hat{y_t}$ and $\hat{e_t}$.} % Figure adapted from \cite{BERTSCHINGER.2006}
					\label{fig:SystemAndEnv}
				\end{figure}

			\noindent
			\cite{BERTSCHINGER.2006} defines that a system is informationally closed when information flow from the environment to the system is zero.

				\begin{equation}
				J_{t}(E \rightarrow Y )=0
				\label{eq:informationflow2}
				\end{equation}

			% ------------------------------- trivial case ------------------------------- %
			\noindent
			Information closure (minimising $J_t$) is trivial if the environment and the system are entirely independent of each other.

				\begin{equation}
				\begin{aligned}
				{I(Y_{t+1};E_{t})=0}&&{\Rightarrow}&&{J_{t}(E \rightarrow Y )=0}
				\end{aligned}
				\end{equation}

			% -------------------------------- non-trivial ------------------------------- %
			\noindent
			However, informational closure can be formed non-trivially. In the non-trivial case, even though a system contains (or encodes) information about the environmental dynamics, the system can still be informationally closed. In such cases, the mutual information between the current states of the environment and the future state of the system is larger than zero.

				\begin{equation}
				I(Y_{t+1};E_{t}) > 0
				\end{equation}

			\noindent
			This also implies
				\begin{equation}
					I(Y_{t+1};Y_{t})-I(Y_{t+1};Y_{t}|E_{t}) > 0
				\end{equation}

			\noindent
			And, non-trivial informational closure can be defined as
				\begin{align}
%				\label{eq:NTIC}
    				% \left.\begin{array}
    				NTIC_t(E\rightarrow Y) :&=I(Y_{t+1};Y_{t})-I(Y_{t+1};Y_{t}|E_{t}) 
    				\label{eq:NTIC1}\\
    				&=I(Y_{t+1};E_{t})-I(Y_{t+1};E_{t}|Y_{t}) 
    				\label{eq:NTIC2}
    				% \end{array} \right.
				\end{align}

			\noindent
			Hence, maximising $NTIC_t(E\rightarrow Y)$ amounts to
				\begin{equation}
    				\label{eq:nticObjective}
    				\begin{aligned}
    				& \text{maximising} & { } & I(Y_{t+1};Y_{t}) & { } & \text{and} \\
    				& \text{minimising} & { } & I(Y_{t+1};Y_{t}|E_{t}) & { }
    				\end{aligned}
				\end{equation}

			\noindent
			One can also maximise $NTIC_t(E\rightarrow Y)$ by 
				\begin{equation}
    				\label{eq:nticObjective2}
    				\begin{aligned}
    				& \text{maximising} & { } & I(Y_{t+1};E_{t}) & { } & \text{and} \\
    				& \text{minimising} & { } & I(Y_{t+1};E_{t}|Y_{t}) & { }
    				\end{aligned}
				\end{equation}

			\noindent
			This implies the system contains in itself all the information about its own future and the self-predictive information contains the information about the environment.

			\subsection{Informational Closure Does not Imply Causality}\label{sec:causality}
                A surprising (but well known; see ???) result from the definition of information flow $J_t(E\rightarrow Y)$ (Eq. \ref{eq:InformationFlow}) is that information flow does not indicate causal dependency from $E_t$ to $Y_{t+1}$ and from $Y_t$ to $Y_{t+1}$. Here we consider two scenarios, called "modelling" and "passive adaptation", which were already noted by \citet{BERTSCHINGER.2006}. In both scenarios, a process can form positive NTIC ($NTIC(E\rightarrow Y)> 0$) and informational closure ($J(E\rightarrow Y)=0$), however, via different causal dependencies.

                In the scenario of \textit{modelling}, to achieve positive NTIC and informational closure, a system can internalise and synchronise with the dynamics of the environment, e.g., model the environment. In this case, the future internal state $Y_{t+1}$ of the system is driven by the current internal state $Y_t$ and the system still retains mutual information with the environment. Having high degrees of NTIC then entails having high predictive power about the environment. This gives biological agents functional and evolutionary advantages.

                In the scenario of \textit{passive adaptation}, counterintuitively, the future system states ($Y_{t+1}$) are  entirely driven by the current environment states ($E_t$) but can also have positive NTIC and informational closure. This happens under the condition that the sensory process $\hat{e}_t$ is deterministic and the system simply copies the sensory values. The system is then a copy of another informationally closed process ($\hat{e}_t$) and is therefore closed itself. At the same time it has mutual information with the process that it is copying.

                However, in most of the realistic cases, the environment is partially observable to the system and, thus, the sensory process is usually not deterministic. Therefore, it is difficult for the system to be informationally closed and have higher NTIC.
                More importantly, we argue in Appendix \ref{sec:appendix} that whenever the environment has itself more predictable dynamics than the observations, there is potential for a process to achieve higher NTIC by modelling the environment than by copying the observations.

                We will see that both scenarios are relevant to ICT in the following sections.

% ============================================================================ %
%                            Neural coarse-graining                            %
% ============================================================================ %
	\section{Coarse-graining in the Neural System} \label{sec:Neural coarse-graining}

		%% stochasticity at microscopic levels
		The formation of NTIC with a highly stochastic process is challenging. NTIC requires the predictability of the system state and is therefore impeded by noise in the system. Information processing at the microscopic scale (the cellular scale) in neural systems suffers from multiple environmental noise sources such as sensor, cellular, electrical, and synaptic noises. For example, neurons exhibit large trial-to-trial variability at the cellular scale, and are subject to thermal fluctuations and other physical noises \citep{faisal2008noise}.

        %% Our argument
		However, it is possible that neural systems form NTIC at certain macroscopic scales through coarse-graining of microscopic neural states. Coarse-graining refers to many-to-one or one-to-one maps which aggregate microscopic states to a macroscopic state. In other words, a number of different micro-states correspond to the same value of the macro-variable \citep{price2007causation}. Coarse-grainings, can therefore form more stable and deterministic state transitions and more often form NTIC processes. For neural systems this means that a microscopically noisy neural system may still give rise to an NTIC process on a more macroscopic scale.

		Indeed, empirical evidence suggests that coarse-graining is a common coding strategy to establish robustness against noise at microscopic scales of the neural system. For instance, the inter-spike intervals of an individual neuron are stochastic. This implies that the state of an individual neuron does not represent stable information. However, the firing rate, i.e. the average spike counts over a given time interval, is more stable and robust against noise such as the variability in inter-spike intervals. Using this temporal coarse-graining strategy, known as rate coding \citep{adrian1926impulses, gerstner2002spiking, maass2001pulsed, panzeri2015neural, stein2005neuronal}, neurons can encode stimulus intensity by increasing or decreasing the firing rate \citep{kandel2000principles}. \citep{stein2005neuronal}. The robustness of the rate coding is a direct consequence of the many to one mapping (i.e., coarse-graining).

		% population code
		Population coding is another example of encoding information through coarse-graining in neural systems. In this coding scheme, information is encoded by activation patterns of a set of neurons (a neuron population). In the population coding scheme, many states for a neuron population map to the same state of macroscopic variables which encode particular informational contents, thereby reducing the influence of noise in individual neurons. That is, stable representations can be formed through coarse-graining the high dimensional state space of a neuron population to a lower dimensional macroscopic state space \citep{kristan1997population, pouget2000information, binder2009encyclopedia, QuianQuiroga2009}. Therefore, individual neuron states (the microscopic scale) are not informative enough about the complete encoded contents at the population scale (the macroscopic scale). Instead, coarse-grained variables are better substrates for stably encoding information and allow the neural system to ignore noisy interactions at the fine-grained scale \citep{Woodward2007-WOOCWA}.

        These two examples show that the known coding schemes can be viewed as coarse-graining, and provide stochastic neural systems with the ability to form more stable and deterministic macroscopic processes for encoding and processing information reliably. We argue that through coarse-graining the neural systems is able to form NTIC processes at macroscopic scales. Based on the merit of coarse-graining in neural systems, we propose a new theory of consciousness in the next section.

% ============================================================================ %
%               A neural coarse graining theory of consciousness               %
% ============================================================================ %
	\section{\acf{OurTheory}}\label{sec:OurTheory}

        In this section, we propose a new theoretical framework of consciousness: \acf{OurTheory}. The main hypothesis is that conscious processes are captured by what we call \textit{C-processes}. We first define C-processes, then state our hypothesis and discuss its implications.

		\begin{figure}[H]
		    \centering
			\includegraphics[width=\textwidth]{WritingMaterials/Fig_FullGraph/FullGraph.pdf}
			\caption{The information flow amounts the universe $X$, the system $S$, the environment of the system $E$, and the coarse-grained process $Y$ of the system $S$. The solid line with a filled arrow from $X_t$ to $X_{t+1}$ represents the microscopic dynamic of the universe. The solid lines with a empty arrow represent directions of coarse-graining. The dashed lines represents virtual dependencies between  two macroscopic variables. The red $Y_t$, $Y_{t+1}$, and the red dashed line in between represents a macroscopic process which forms informational closure at a certain coarse-grained scale.}
			\label{fig:fullgraph}
	   	\end{figure}

	   	% ------------------------------- definition 1 ------------------------------- %
	   	In order to define C-processes we need to define coarse-grainings first. Every coarse-graining is characterised by a function that maps the microscopic process to the coarse-grained macroscopic process. More formally:

	   	\begin{definition}
	   	\linelabel{line:r2q1-1} Given a stochastic process $X$ with state space $\mathcal{X}$, a \emph{coarse-graining of $X$} is a stochastic process $Y$ with state space $\mathcal{Y}$ such that there exists a function \footnote{Functions in the mathematical sense used here are always either one-to-one or many-to-one.} $f_Y:\mathcal{X}\rightarrow \mathcal{Y}$ with $Y_t=f_Y(X_t)$.  
	   	\end{definition}
	   	A more general definition of coarse-grainings that maps temporally extended sequences of the microscopic process to macroscopic states are possible but for this first exposure of our theory the simpler definition above is sufficient.

	   	% ------------------------------- definition 2 ------------------------------- %
        \begin{definition}
        Given a stochastic process $X$ called the universe process, a \emph{C-process} is a coarse-graining $Y$ of $X$ such that the following two conditions are satisfied (see Fig.~\ref{fig:fullgraph}):
        \begin{enumerate}
        \item $Y$ is informationally closed to $X$
        \item there exists a pair $(S,E)$ of coarse-grainings of $X$ such that 
        \begin{itemize}
            \item $Y$ is a coarse-graining of $S$,
            \item the state space $\mathcal{X}$ of $X$ is equal to the Cartesian product of the state spaces $\mathcal{S}$ and $\mathcal{E}$ of processes $S$ and $E$ respectively, formally $\mathcal{X}=\mathcal{S}\times\mathcal{E}$, and 
            \item $Y$ is NTIC to $E$, formally:
        \begin{align}
            NTIC_t(E\rightarrow Y) >0
        \end{align}
        \end{itemize} 
        \end{enumerate}
       \end{definition}

    % ==================================== %
    % martin.biehl: 
    % doesn't work because the information closrue with respect to $X$ implies full closrure with respect to every 
    % other process. This means all conscious processes are fully closed. This contradicts what we say below.
    % Apr 7, 2020 3:23 AM
    % ==================================== %
    %     \begin{definition}
    %     Given a stochastic process $X$ called the universe process, a \emph{C-process} is a coarse-graining $Y$ of $X$ such that the following two conditions are satisfied (see Fig.~\ref{fig:fullgraph}):
    %     \begin{enumerate}
    %     \item $Y$ is informationally closed to $X$
    %     \item there exists a pair $(S,E)$ of coarse-grainings of $X$ called the \textit{system} $S$ and its \textit{environment} $E$  such that 
    %     \begin{itemize}
    %         \item $Y$ is a coarse-graining of $S$,
    %         \item the state space $\mathcal{X}$ of $X$ is equal to the Cartesian product of the state spaces $\mathcal{S}$ and $\mathcal{E}$ of processes $S$ and $E$ respectively, formally $\mathcal{X}=\mathcal{S}\times\mathcal{E}$, and 
    %         \item $E$ is the environment process of $S$ with respect to which $Y$ is maximally NTIC:
    %         \begin{align}
    %         f_E:=\argmax_{\{\E||\E|\leq |\X|\},g:\X \rightarrow \E} NTIC(g(X)\rightarrow Y) \end{align}
    %     %     $Y$ is NTIC to $E$, formally:
    %     % \begin{align}
    %     %     NTIC_t(E\rightarrow Y) >0
    %     % \end{align}
    %     \end{itemize} 
    %     \end{enumerate}
    %   \end{definition}
    % ====================================%

        \noindent Note that, here we applied the same definitions of information flow (Eq.~\ref{eq:InformationFlow}) \linelabel{line:IC2CG}
        \begin{align}
            J_{t}(E \rightarrow Y )=I(Y_{t+1};E_t|Y_t)
        \end{align}
            to the system-environment dependency
            and the micro-macro scale dependency
        \begin{align}
            J_{t}(X \rightarrow Y )=I(Y_{t+1};X_t|Y_t)
        \end{align}  
        even though the Bayesian graphs are different in the two scenarios. Both of these settings have already been used in the literature \citep[see][]{BERTSCHINGER.2006, PFANTE.2014}).\\\\

        With the two definitions we can state the main hypothesis of \ac{OurTheory}
        \begin{hypothesis}
        A process $Y$ is conscious if and only if it is a C-process of some process $X$. Also the content of consciousness $C_t^{Content}$ at time $t$ is the state $y_t$ of the C-process at time $t$ and the level of consciousness $C_t^{Level}$ is the degree of NTIC of the process to the environment i.e. $NTIC_t(E\rightarrow Y)$:
        \begin{align}
            C_t^{Content} &= y_t \label{eq:cContent}\\
            C_t^{Level} &= NTIC_t(E\rightarrow Y) \label{eq:cLevel}
        \end{align}
        \end{hypothesis}

        % ====================================
        % Refinement construction approach
        % ====================================
        % \begin{hypothesis}
        %         For any coarse-graining $Y$ of a stochastic process $X$, a refinement $S$ of $Y$ and the environment $E$ of $S$.  
        %         The \emph{content $C_t^{Content}(Y)$ of the consciousness of $Y$ at time $t$} is the state $y_t$ of $Y$ at time $t$ and \emph{the level $C_t^{Level}(Y)$ of the consciousness of $Y$} is the degree of NTIC of the C-process to the environment minus the information flow from the microscopic process $X$ to $Y$:
        %         \begin{align}
        %             C_t^{Content} &= y_t \label{eq:cContent}\\
        %             C_t^{Level} &= NTIC_t(E\rightarrow Y)-J_t(X \rightarrow Y) \label{eq:cLevel}
        %         \end{align}
        %         \end{hypothesis}
        % ====================================

        A concrete example in the context of neuroscience is that $X$ represents the microscopic scale of the universe, $S$ a cellular scale process in the neural system, $Y$ a more macroscopic process of the neural system coarse-grained from the cellular scale process $S$, and $E$ the environment which the cellular level process $S$ interacts with. \linelabel{line:neu-env} The environment $E$ may include other processes in the neural system, the sensors for perception and interoception, and external physical worlds.\\

        Based on the hypothesis, \ac{OurTheory} leads to five core implications: 
        \begin{description}
            \item[Implication 1.] 
            Consciousness is information. Here, "informative" refers to the resolution of uncertainty. Being in a certain conscious state rules out other possible conscious states. Therefore, every conscious percept resolves some amount of uncertainty and provides information. \\ 
            This implication is also in agreement with the "axiom" of \textit{information} in Integrated Information Theory (IIT 3.0) which claims that \textquote[{\citealp[P.~2]{oizumi2014phenomenology}}]{\dots an experience of pure darkness is what it is by differing, in its particular way, from an immense number of other possible experiences\dots}

            \item[Implication 2.] 
            Consciousness is associated with physical substrates and the self-information of the conscious percept is equal to the self-information of the corresponding physical event. This is a direct implication from our hypothesis that every conscious percept $C_t^{Content}$ corresponds to a physical event $y_t$.

            \item[Implication 3.] 
            Conscious processes are self-determining. This is a direct implication of the requirement that $Y$ is informationally closed with respect to $X$. To be informationally closed with respect to $X$, no coarse-graining knows anything about the conscious process' future that the conscious process does not know itself. This self-determining characteristics is also in line with our daily life conscious experience which often shows stability and continuity and is ignorant of the stochasticity (e.g., noise) of the cellular scales.

            \item[Implication 4.] 
            Conscious processes encode the environmental influence on itself. This is due to the non-triviality of the informational closure of $Y$ to $E$. At the same time all of this information is known to the conscious processes themselves since they are informationally closed with respect to their environments. This also suggests that conscious processes can model the environmental influence without knowing more information from the environment.

            \item[Implication 5.]
            Conscious processes can encode environmental information (by forming NTIC), however, be ignorant to part of the information of more microscopic processes (from Implication 3 and 4). This is in line with our conscious experience that information that every conscious percept provides represent rich and structured environmental states without involving all the information about microscopic activities.
        \end{description}

		% ---------------------------------------------------------------------------- %
        %        Level of Consciousness correlates degrees of NTIC of a process        %
        % ---------------------------------------------------------------------------- %
	    \subsection{Level of Consciousness is Equal to the Degree of NTIC of a Process}\label{sec:cl}
            According to Eq.~\ref{eq:nticObjective}, \ac{OurTheory} implies that conscious levels are determined by two quantities.

            First, to form a high level of NTIC, one can increase the mutual information $I(Y_{t+1};Y_{t})$ between the current internal state $Y_t$ and the future internal state $Y_{t+1}$. In other words, conscious levels are associated with the degree of self-predictive information \citep{bialek2001predictability}. This mutual information term can be further decomposed to two information entropy quantities:

            \begin{equation}
            \label{eq:SelfEntropy}
            I(Y_{t+1};Y_{t}) = H(Y_{t+1}) - H(Y_{t+1}|Y_t)
            \end{equation}

            This implies that a highly NTIC process must have rich dynamics with self-predictability over time. Another implication is that complex systems can potentially attain higher levels of consciousness due to the larger information capacities needed to attain high mutual information. This outcome is consistent with the common intuition that conscious levels are often associated with the degree of complexity of a system.

    	    Second, one can minimise the conditional mutual information $I(Y_{t+1};Y_{t}|E_{t})$ to increase the level of NTIC. If the mutual information term $I(Y_{t+1};Y_t)$ is supposed to stay large, this quantity suggests that conscious level increases with the amount of information about the environment state $E_t$ that the NTIC process encodes in its own state $Y_t$ and $Y_{t+1}$. In other words, $Y_t$ should not contain more information about $Y_{t+1}$ than $E_t$. An important implication is that agents interacting with a complex environment have the chance to build a higher level of NTIC within their systems than those living in a simple environment. In other words, the level of consciousness is associated with environmental complexity.

    	    %% not monotonic
    		\linelabel{line:non-monotonic}It is important to note that NTIC can be a non-monotonic function of the scale of coarse-graining. At the finest scale we consider the whole universe $X$ as the process $Y$. then, since $Y$ is a coarse graining of $S$ we have $Y=S=X$. In this case the environment $E$ corresponding to the universe seen as a system is the constant coarse-graining\footnote{Recall that, for a system with state space $\S$ the environment state space $\E$ must be such that $\X=\S \times \E$. If $\S =\X$ then we need $\E$ with $\X \times \E=\X$ such that $\E$ must be a singleton set. All coarse-grainings mapping $\X$ to a singleton set are constant over $\X$.} and therefore the mutual information $I(E_t;Y_{t+1})$ and the transfer entropy $I(Y_{t+1};E_t|Y_t)$ are zero. Then NTIC of the universe with respect to its environment is zero and $X$ can never be a C-process.

    		If we now increase the scale of $Y$, this allows $S$ to also reduce in scale and therefore $E$ can become more and more fine-grained. This means that the mutual information $I(E_t;Y_{t+1})$ between $E$ and $Y$ can at least potentially become positive. Up to the point where $E$ accounts for half of the bits of $X$ and $S$ for the other half the upper bound of the mutual information $I(E_t;Y_{t+1})$ achieved when $Y=S$ increases. Even further refining $E$ then again leads to a reduction of the upper bound of $I(E_t;Y_{t+1})$.
            % ==================================== %
            %    WHY DO WE NEED the FUCKING $Y$    %
            % ==================================== %
            % 2020/04/07 01:57
    		% Why does Definition 2 have a fucking $Y$? It seems the whole thing could be defined wihtout referring to the underlying $S$ and directly use $Y$'s own environment (together with which it corresponds to $X$). We would always have a refinement of $Y$ that could be called $S$ to gether with its environment. But this is alwyas the case so it doesn't restrict the definition of $Y$ at all. (Especially if we can just choose $S=Y$). The only thing that makes the current definition of $Y$ more general is that it may be that $Y$ is only closed with respect to a coarsening of its own environment and not to its own environment itself. This is a waekening of the restriction compared to requiring that $Y$ be NTIC with respect to its own environment and IC with respect to $X$. Actually no, since $Y$ is IC with respect $X$ it is IC with respect to everything including to any coarsening of its own environment. Since the mutual information between $Y$ and its environment is greater or equal to the mutual information between $Y$ and a coarsening of its environment $Y$ must also be NTIC with respect to its own environment. 
    		% ==================================== %

    		On the other extreme, when $E=X$ the system state space must be the singleton set and NTIC from $E$ to $Y$ must again be zero. Therefore, processes at intermediate scales of coarse-graining can form higher degrees of NTIC than at the most microscopic or macroscopic scales (Fig.~\ref{fig:LevelOfConsciousness}). \ac{OurTheory} indicates that human consciousness occurs at the scale of coarse-graining where higher NTIC is formed within the neural system.

    		\begin{figure}[H]				
        		\includegraphics[width=\textwidth]{Submission/Frontiers/nonMono.pdf}
        		\caption{A non-monotonic relationship between scale of coarse-graining and level of consciousness.}
        		\label{fig:LevelOfConsciousness}
    		\end{figure}

		\subsection{Conscious Contents Corresponding to States of a NTIC Process}\label{sec:cc}
    	    % richness 
    		\ac{OurTheory} proposes that conscious contents correspond to the states of NTIC processes (Eq.~\ref{eq:cContent}). This implies that the size of the state space of an NTIC process is associated with the richness of conscious contents that the process can potentially have. Therefore, a complex NTIC process with a high dimensional state space can have richer conscious experience than a simple NTIC process can have. This outcome is consistent with the intuition that the richness of conscious contents is associated with the complexity of a system.

    		% CC doesn't include all microscopic states
    		As mentioned above, informational closure can happen between scales of coarse-graining within a single system. Thus, a macroscopic NTIC process can be ignorant to its microscopic states. \ac{OurTheory} argues that human conscious contents do not reflect cellular scale activity because the conscious process which corresponds to a macroscopic NTIC process is informationally closed to the cellular scale in the human neural system. Further more, since NTIC processes are informationally closed, each of them can be considered as a reality. In the extreme case, when the information flow from its microscopic processes and environment to the informationally closed process is completely zero (Eq.~\ref{eq:informationflow2}), the future states of the process is only determined by its past states.

    		\linelabel{line:encode}Importantly, NTIC processes internalise the environmental dynamics in its states \citep[also see P.~4][]{BERTSCHINGER.2006}. This suggests that a NTIC process can be considered as a process that models the environmental dynamics.This implication fits well with some theories of consciousness (for example, world simulation metaphor ~\citep{revonsuo2006inner}). Note that \ac{OurTheory} doesn't assume that generative models are necessary for consciousness. The implication is a natural result of processes with NTIC.

            Finally, a coarse-graining can be a many to one map from microscopic to macroscopic states and \ac{OurTheory} proposes that conscious contents $C^{Content}$ is the state of the NTIC process $Y$. Therefore, \ac{OurTheory} implies multiple realisation thesis of consciousness \citep{putnam1967psychological,bechtel1999multiple} which suggests that different physical implementations could map to the same conscious experience.

		% ---------------------------------------------------------------------------- %
        %             Reconciling the levels and contents of consciousness             %
        % ---------------------------------------------------------------------------- %
	    \subsection{Reconciling the Levels and Contents of Consciousness}\label{sec:reconcile}
    	    While it is useful to distinguish the notion of the levels and contents of consciousness, whether they can be clearly dissociated has been a matter of debate \citep{bayne2016there, Fazekas2016}. In \ac{OurTheory}, conscious levels and conscious contents are just two different properties of NTIC processes, and it therefore naturally reconciles the two aspects of consciousness. In an NTIC process with a large state space, conscious contents should also consist of rich and high dimensional information. Therefore, this framework integrates the levels and the contents of consciousness in a coherent fashion by providing explicit formal definitions of the two notions.

    	    According to Sec.~\ref{sec:cl} and Sec.~\ref{sec:cc}, an important implication from \ac{OurTheory} is that both conscious levels and conscious contents are associated with the state space of an NTIC process $Y$. A large state space of $Y$ contributes conscious levels through the mutual information $I(Y_{t+1};Y_{t})$ and also contributes richer conscious contents by providing more possible states of conscious processes. 
    	    \ac{OurTheory} therefore explains why, in normal physiological states, conscious levels and conscious contents are often positively correlated \citep{laureys2005neural}. This implication is also in line with the intuition in which consciousness is often associated with complex systems.

    % ============================================================================ %
    %                    Conscious versus Unconscious Processing                   %
    % ============================================================================ %
	\section{Conscious Versus Unconscious Processing}\label{sec:Conscious versus Unconscious Processing}
	    In this section, we show how \ac{OurTheory} can explain and make predictions about what processes are more conscious than others. \ac{OurTheory} is constructed using information theory. Therefore, \ac{OurTheory} can provide predictions based on the mathematical definitions.

        % ---------------------------------------------------------------------------- %
        %                            Unconscious Processing                            %
        % ---------------------------------------------------------------------------- %
        \subsection{Unconscious Processing}
            In this section we highlight two scenarios in which ICT predicts that processes remain unconscious. 
            % e or  decrease the degree of NTIC of a process, thereby making the process less conscious or unconscious.

            \subsubsection*{Processes that are not Informationally Closed}\label{sec:reflexive}
                % ------------------------------------
                % This section should follow:
                % 1. Processes that are not informationally closed is unconscious
                % 2. Some informationally not closed processes are passively driven by the environment (some examples)
                % 3. This may give readers the impression that processes passively driven by the environment are all not informationally closed
                % 4. This is a wrong impression because in a specific circumstance which the sensory processes are deterministic, passive adaptation can be informational closed.
                % 5. However, 4 is difficult to get because in realistic cases, the environment is often partially observable and this leads to sensory processes being non-deterministic. Therefore, ICT predict that in most of the realistic cases, passive adaptation (e.g., feed-forward network, video camera) is unconscious.
                % ------------------------------------
                \linelabel{line:notIC} 
                % To maximise NTIC, one can minimise the information flow $J_{t}(E \rightarrow Y )$ from the environment $E$ to the process $Y$ (Eq.~\ref{eq:InformationFlow},\ref{eq:nticObjective2}). When the information flow is zero, the process is informationally closed with respect to the environment. 
                The first scenario is built upon the assumption that sensor values are non-deterministic and the process dynamics are passively driven by the environmental inputs. Such processes cannot be informationally closed and are therefore unconscious.

                % When the sensor processes are not deterministic, if processes dynamics is passively driven by the environmental inputs the process is not informationally closed and, thus, unconscious.

                Reflexive behaviours \citep{casali2013theoretically} can be considered as an example of this scenario. In \ac{OurTheory}, if we can view reflexive behaviours as situations in which the internal state $Y_t$, which triggers reflexive action, is determined by the environment state $E_{t-1}$, overruling the influences from its own past $Y_{t-1}$. Such interpretation of reflexive behaviour from the viewpoint of \ac{OurTheory} naturally explains why reflexes involve less or no conscious experience of external stimuli.

        		\begin{figure}[H]
        			\includegraphics[width=0.8\textwidth]{Submission/Frontiers/Reflexive.pdf}
        			\caption{
        			    A diagram depicting the information flow in reflexive behaviours (shown by the red nodes and arrows) happening through the interaction between a process $Y$ and its environment $E$. In such situations, the internal state $Y_t$ is mostly dependent on the environment state $E_{t-1}$ but less on its past state $Y_{t-1}$. Therefore, the process $Y$ is not informational closed. As a  consequence, $Y$ is unable to form high NTIC and, therefore, remain less conscious or unconscious (see Fig.~\ref{fig:reflexive}). 
        			    }
        			\label{fig:reflexive}
        		\end{figure}

                The same principle can be applied to interpret blindsight \citep{humphrey1999history, humphrey1974vision, Humphrey1970} and procedural memory \citep{doyon2009contributions, ashby2010cortical} which are often considered as unconscious processes.
        		\linelabel{line:r2q6}Blindsight patients are able to track objects, avoid obstacles, and make above chance-level visual judgements with degraded or missing visual experience. However, in some cases, they may still preserve some forms of conscious experience (see \cite{overgaard2011visual, mazzi2016blind}). We argue that blindsight-guided actions are a result of 
        		% action outputs of blindsight can be directly guided by sensory inputs through 
        		stimulus-response mapping. The corresponding neural circuits are driven passively and therefore are not informationally closed. According to \ac{OurTheory} we therefore have no conscious experience of the visual stimuli.

        		Similarly, for procedural memory, the state transitions of corresponding neural circuits determining the action sequences largely depend on  sensory inputs. This prevents the neural processes of procedural memory from informational closure and being conscious.
        		\ac{OurTheory} also offers an interpretation as to why patients with visual apperceptive agnosia \citep{james2003ventral} can perform online motor controls without visual awareness of action targets \citep{10.3389/fneur.2014.00255}.

                Note that, not all processes that are driven by the environment (passive adaptation) are unconscious. When the sensor processes are deterministic, a system can still have positive NTIC and achieve informational closure via passive adaptation. Therefore, some passive system (for example pure feedforward networks) can potentially produce consciousness.

                For agents such as human beings, the environment is often informationally rich but only partially observable in such a way that the current sensory inputs are insufficient to predict the next and to form deterministic sensor processes. In this circumstance, the system cannot become informationally closed by passive adaptation (e.g., simply copying the sensory values to the system). ICT predicts that, in most of the realistic cases, processes with passive adaptation are unconscious. On the other hand, networks with recurrent loops employ information stored in their own past states have the potential to achieve higher NTIC by modelling the environment. If it turns out to be true that for every pure feed-forward network there are non-feed-forward systems achieving higher NTIC then ICT predicts that the latter systems achieve higher levels of consciousness. This implication would coincide with theories of consciousness emphasising the importance of recurrent circuits to consciousness \citep{lamme2006towards, edelman1992bright, tononi2008neural}.

                % ==== end ===

                % Finally, with a caveat, when the sensory dynamics is deterministic, a system can still have positive NTIC and achieve informational closure by simply copying the sensory values to its internal state (i.e. passive adaptation, see Sec.~\ref{sec:causality}). ICT will predict that, in this specific scenario, the systems can have some levels of consciousness. Although this specific scenario is uncommon in realistic circumstance an important implication is that pure feedforward networks can potentially produce consciousness. The level of consciousness for these networks is bounded by the sensory channel capacity and depth of the network. On the other hand, networks with recurrent loops employ information stored in their own past states have the potential to achieve higher NTIC by modelling the environment. If it turns out to be true that for every pure feed-forward network there are non-feed-forward systems achieving higher NTIC then ICT predicts that the latter systems achieve higher levels of consciousness.

                % Namely, if the future state of a process depends primarily on the sensor value but receives little influence from its past state.

            %  The first scenario is that of a system partially observing the environment in such a way that the current observation is insufficient to predict the next. In this situation we can show that a system passively processing the observations (including copying them) achieves lower values of NTIC and therefore lower levels of consciousness than a system that is able to accurately predict the observations.

            % \paragraph{@ Acer's version}
            % Based on Eq.~\ref{eq:nticObjective2}), to maximise NTIC, a process can minimise the information flow from the environment to the process $J_{t}(E \rightarrow Y ) := I(Y_{t+1};E_{t}|Y_{t})$ (Eq.~\ref{eq:InformationFlow}). For agents such as human being, who always face informationally rich environments, the environment is usually partially observable with limited sensory capacities. This leads to the sensor values are not deterministic (or close to it). In such cases, if the state of a internal process is mainly driven by sensory inputs, the information flow from the environment to the process is considered to be large and the process is not informationally closed. 
            % First, to maximise NTIC, a process can minimise the information flow from the environment to the process $J_{t}(E \rightarrow Y ) := I(Y_{t+1};E_{t}|Y_{t})$ (Eq.~\ref{eq:InformationFlow}). Here, we illustrate two distinctive scenarios which shade light of the relationship between informational closure and consciousness. 
            % \paragraph{@ Martin's version}
            %Is there a (generative) epsilon machine that achieves higher NTIC than copying? If yes, under what conditions (e.g. reversibility of the underlying environment process?) on the sensor process does it exist?
            %Is Bertschinger's system better than copying?

            % ------------------------------------
            % \ac{OurTheory} leads to some surprisingly simple conscious processes. Consider the case already noted by \citet{BERTSCHINGER.2006} where the environment process is deterministic (or close to it) and the system just copies the environment process. Then the system can achieve positive values of NTIC with respect to the environment. ICT states that such processes are conscious. 
            % ------------------------------------

            % (We assume Transfer entropy from environment to system is equal to transfer entropy from sensor to system by definition).

            % However, in more realistic cases the system cannot copy the entire environment state because it observes only part of it. 
            % and/or because it is itself not large enough to contain a copy of the environment.

            % We show in Appendix \ref{sec:appendix} that whenever the environment has itself more predictable dynamics than the observations there is potential for a predictive process to achieve higher NTIC than a copying system or any system that only processes its last observations without taking into account other internal memory. Similar to \citet{BERTSCHINGER.2006} we will refer to systems of the latter type as \textit{passive}. We will call all other systems that achieve positive NTIC \textit{predictive}. Note that the predictive process might need to be more complex than the system that only copies its observations.

            % We tentatively assume below that for a feed-forward network which can be seen as implementing a function from an extended history of past observations to the system state there also always exist systems that achieve higher NTIC. This remains to be verified.

            % Notes: 
            % -put complicated stuff into appendix
            % -make clear notion of passive system vs predictive system.
            % This shows that for 
            % \begin{align}
            %  I(S_{t+1}; f(E_t)|S_t)=I(g\circ f(E_t);f(E_t)|f(E_{t-1}))=H(f(E_t)|f(E_{t-1}).
            % \end{align}

            % In such situations ITC predicts that higher levels of consciousness are achieved if the system can anticipate the next observed values which requires a higher internal self-determination than that of the copies of the observations.

            % Another way to put this is the higher the difference in predictability of the sensor values and the entrie environment state the more potential is there to increase the level of consciousness for the system by modelling the underlying hidden Markov process i.e.\ environment.

            % Also assume that the part that the system observes is not itself deterministic or close to it but that the underlying environment is.

            % A system with a high NTIC in such situations must 
            % Then the system can decrease the transfer entropy from the environment to itself by anticipating the next observation. This is due to the fact that the transfer entropy from the environment to the system is equal to the transfer entropy from the observations to the system.

            % does not have access to the entire environment state and instead observes only part of it. This part is rarely close to deterministic even if the entire environment is deterministic. In order to achieve informational closure with respect to the entire environment the system has to achieve closure exactly with respect to its sensor values.

            % we could copy it to get closure. In realistic situations we are not sure whether the environment is informationally closed but even if we assume it is we can confidently assert that we don't observe it completely and that the part we do observe (our sensory values) are not themselves closed. For example, the images on our retina are insufficient to predict their successors without imagining the outside world behind them. This imagining however requires internal processing like recurrent connections.

           	% \textit{=== End of the modification === }\\

                % The higher the transfer entropy from environment to the system the lower NTIC (Eq.~\ref{eq:nticObjective2}) and therefore, according to ICT, the lower the level of consciousness. Under the assumptions above,

                % high transfer entropies occur when the current state of a process depends primarily on the environment state (see Fig.~\ref{fig:reflexive}), but receives little influence from its past state.

            %     Let us now look at some examples of this scenario.

            %   The first are reflexive behaviours \citep{casali2013theoretically}. 
            % %   can be considered an example of this scenario. 
            %   In \ac{OurTheory}, if we can view reflexive behaviours as situations in which the internal state $Y_t$, which triggers reflexive action, is determined by the environment state $E_{t-1}$ overruling the influences from its own past $Y_{t-1}$.  Such interpretation of reflexive behaviour from the viewpoint of \ac{OurTheory} naturally explains why we are less conscious of external stimuli causing reflexes than of others.

             	% reflexive behaviour

        %         their conscious levels are bounded by their sensory channel capacity.

        %         % \paragraph{@ Acer's revision}
        %         % However, with a caveat, informational closure can also be achieve by just copying the environment process in some certain cases. Consider the case already noted by \citet{BERTSCHINGER.2006} where the environment process is deterministic (or close to it) and full observe.  the system just copies the environment process. Then the system achieves a high value of NTIC with respect to the environment. ICT states that such processes are conscious.

        %         % \paragraph{@ Orignal}

        % 		A crucial implication of ICT is that a pure feedforward network in the most of

        % 		cannot produce consciousness because NTIC requires a form of memory I(Y t+1 ; Y t ) . Without memory, a network’s current state is entirely driven by the input the network receives from the environment without any inﬂuence from its own past states. Therefore, such a network is incapable of forming an NTIC. In contrast, a network with recurrent loops can maintain information about its own past states. This forms an information channel between the past and the future states of the network and, thus, makes the network capable of being informationally closed. This result coincides with theories of consciousness emphasising the importance of recurrent circuits to consciousness (Edelman, 1992; Lamme, 2006; Tononi & Koch, 2008).
        % 		than the because NTIC requires a form of memory $I(Y_{t+1};Y_{t})$. Without memory, a network's current state is entirely driven by the input the network receives from the environment without any influence from its own past states. Therefore, such a network is incapable of forming an NTIC. In contrast,

            \subsubsection*{Processes are Trivially Closed}
                The second scenario is that when encoded information in a process is trivial, i.e. no mutual information between the process states and its environment states $I(Y_{t+1};E_{t})$ (Eq.~\ref{eq:nticObjective2}), this leads to non-positive NTIC. In such cases, the process is considered to be unconscious. This implies that an isolated process which is informationally closed is insufficient to be conscious. 
                This mathematical property of \ac{OurTheory} provides a natural and intuitive (but only partial, see the current challenge in  Sec.~\ref{sec:Limitation and Future work}) solution to the boundary and the individuality problem of consciousness
                    \footnote{The boundary problem of consciousness refers to identifying physical boundaries of conscious processes and the individuality problem of consciousness refers to identifying individual consciousnesses in the universe.}
                \citep{Raymont2006-RAYUOC}. Consider a NTIC process $Y$ and an isolated informationally closed process $\hat{Y}$ with only trivial information. Adding $\hat{Y}$ to $Y$ can still maintain informational closure but does not increase non-trivial information, i.e.,  consciousness is unaffected.

    			\begin{equation}
    			    \begin{aligned}
                        I(Y,\hat{Y};E) & = H(Y,\hat{Y}) - H(Y,\hat{Y}|E) \\
                                       & = H(Y) + H(\hat{Y}|Y) - (H(Y|E)+H(\hat{Y}|Y,E)) \\
                                       & = H(Y) + H(\hat{Y}) - (H(Y|E)+H(\hat{Y})) \\
                                       & = H(Y) - H(Y|E)\\
                                       & = I(Y;E)				
    				\end{aligned}
    			\end{equation}

                This implies that isolated processes with trivial information do not contribute consciousness and should be considered being outside the information boundary of the conscious processing (for more details of the boundary detection procedure, see \cite{krakauer2014information}). 
                This property also implies that consciousnesses do not emerge from just aggregating informationally closed (isolated) processes which contain trivial information.%\footnote{However, we do not exclude the possibility that triviality of encoded information is just a proxy of other quantities which more directly associate with consciousness. We point out that this is a limitation of the current version of \ac{OurTheory} in Sec.~\ref{sec:Limitation and Future work}.}

        % ---------------------------------------------------------------------------- %
        %                             Conscious Processing                             %
        % ---------------------------------------------------------------------------- %
		\subsection{Conscious Processing}\label{sec:conscious processing}
		    According to \ac{OurTheory}, we claim that any process, system, or cognitive function which involves any NTIC process should be accompanied by conscious experience.

		    Previous consciousness research has identified a number of diverse cognitive processes often accompanied by conscious experience. \ac{OurTheory} provides an integrated account for the reason why these processes involve conscious experience. As mentioned above, an NTIC process can be seen as an internal simulation engine for the agent-environmental interactions \citep{BERTSCHINGER.2006}. Therefore, information encoded in NTIC processes is essential for several cognitive processes.

		    One of the most valuable information is the predictions about the environmental states. Cognitive functions requiring agent-scale environmental predictions are likely to recruit NTIC processes and therefore accompanied by conscious experience, for example planning and achieving long term goals.

            Second, as a simulation engine, with a given initial state, an NTIC process can self-evolve and simulate the environmental transitions. Cognitive functions involving simulations are expected to involve NTIC processes. Consequently, mental simulation, imagination, computing alternative realities, and generating counterfactuals often come with conscious experience.

    	    Third, as an informationally closed system, an NTIC process can still provide environmental information without new sensory inputs. This is crucial for many types of off-line processing. Therefore, in contrast to reflexive-like behaviours mentioned above (Sec.~\ref{sec:reflexive}), behaviours requiring off-line computations \citep{milner1999paradoxical, himmelbach2005dorsal,revol2003pointing} often involve conscious experience.

    	    Finally, for agents adapting to complex environments (e.g., human being), any state of the NTIC process can be seen as an integration of high dimensional information. To accurately encode information about the complex environmental states and transitions, the NTIC process requires knowledge about the complex causal dependencies involved in the environment. Therefore, cognitive functions requiring large scale integration are likely to involve NTIC processes and accompanied by conscious experience. % Furthermore, the high dimensional integrated information is also critical for an agent to learn new behaviours to achieve flexibility and the strategic behavioural controls and decision-making.

    	    Note that many of the claims above are compatible with several theories of consciousness which highlight the connection between consciousness and internal simulation, predictive mechanism, or generative models inside a system (e.g. world simulation metaphor ~\citep{revonsuo2006inner}, predictive processing and Bayesian brain~\citep{clark_2013,Hohwy2013,seth2014predictive}, generative model and information generation~\citep{kanai_chang_yu_de_abril_biehl_guttenberg_2019}). Instead of relating functional or mechanistic aspects of a system to consciousness, \ac{OurTheory} captures common informational properties underlying those cognitive functions associated with consciousness. As such, \ac{OurTheory} does not assume any functionalist perspective of consciousness, which associate specific functions to consciousness.  That is to say, since \ac{OurTheory} associates information  with consciousness, functional features accompanied by consciousness are collateral consequences of neural systems utilising NTIC processes for adaptive functions.

    	    In sum, we argue that cognitive functions involving the NTIC process are inevitably accompanied by consciousness. Having an NTIC process is potentially an effective approach to increase fitness in the evolution. It is likely that biological creatures evolve NTIC processes at some point in the evolution. Due to the fundamental relation between information and consciousness, biological creatures also evolve different degrees of consciousness depending on the physical scales and the complexity of the environments they adapt to.

    	    \ac{OurTheory} starts with a non-functional hypothesis, however, it accounts for the association between  functional and consciousness. \ac{OurTheory} further demonstrates remarkable explanatory power for various findings of conscious and unconscious processing.

    % % ============================================================================ %
    % %                        Comparison with other theories                        %
    % % ============================================================================ %
    \section{Comparison with Other Relevant Theories of Consciousness}\label{sec:Comparison with other theories}
    In this section, we compare \ac{OurTheory} with other relevant theories of consciousness.

        % ---------------------------------------------------------------------------- %
        %        Theories about multilevel views on consciousness and cognition        %
        % ---------------------------------------------------------------------------- %
        \subsection{Multilevel Views on Consciousness and Cognition}\label{sec:MultiLevelView}
    		\ac{OurTheory} proposes that conscious processes can occur at any scale of coarse-graining which forms NTIC within a system. This suggests that the scale of coarse-graining is critical for searching and identifying the information corresponding to consciousness. A few versions of multilevel views on consciousness have previously been (explicitly or implicitly) proposed. To our knowledge, Pennartz's neurorepresentational theory (also called Neurorepresentationalism,~\citep{pennartz2018consciousness,pennartz2015brain}) is the proposal closest to the multilevel view of \ac{OurTheory}. Similar to Neurorepresentationalism, the concept of levels in \ac{OurTheory} is also relevant to Marr's level of analysis~\citep{marr1982vision, pennartz2015brain, pennartz2018consciousness}. 
    		However, \ac{OurTheory} suggests that coarse-graining is necessary only when the microscopic processes are stochastic (e.g. the neural system). An NTIC process can be formed in a noise-free deterministic system without coarse-graining. According to \ac{OurTheory}, this NTIC process is sufficient to be conscious. 
    		Another fundamental difference between \ac{OurTheory} and Neurorepresentationalism is that Neurorepresentationalism takes functionalist perspective and suggests consciousness should serve high-level world-modelling and makes a best guess about the interaction between the body and the environment. 
    		%Neurorepresentationalism also suggests conscious experience is associated with integrated representation for multimodal and situational information.
    		However, \ac{OurTheory} is grounded by non-functional informational hypothesis. Therefore, \ac{OurTheory} provides a more fundamental explanation for the scale problem of consciousness.

    		Another well-known proposal based on multilevel views is the Intermediate Level Theory of Consciousness \citep[ILT]{prinz2007intermediate, jackendoff1987consciousness}. ILT proposes that conscious experience is only associated with neural representations at intermediate \textbf{levels of the sensory processing hierarchy} (e.g., the 2.5D representation of visual processing) rather than lower (e.g., pixel) or higher (e.g., abstract) levels of the sensory hierarchy.

    		Here, we want to make clear that the "level" in \ac{OurTheory} refers to the \textbf{scales of coarse-graining} instead of the "level" for the cortical anatomy or sensory processing. It is important to note that the coarse-graining direction is an orthogonal dimension irrespective of the level of anatomy or the level of information processing hierarchy in the neural system (see Fig.~\ref{fig:hierarchy}). Because ILT focuses on the levels of the sensory processing hierarchy and \ac{OurTheory} focus on informational closure among the levels of coarse-graining, the two theories are fundamentally different.

    		\begin{figure}[H]
    			\includegraphics[width=\textwidth]{WritingMaterials/Fig_SeperationOfCGandCortHierachy/SeperationOfCGandCortHierachy.pdf}
				\caption{The distinction between the level of coarse-graining and the level of the cortical hierarchy. $X$ and $Y$ represent the microscopic and macroscopic coarse-grained  variables, respectively. $X^0$ represents the microscopic states at the upstream of the cortical hierarchy. The red empty arrows represents the directions of coarse-graining and the blue arrows represent the directions of the physical dependencies in the cortical hierarchy from the upstream to the downstream. (Some variables and dependencies are omitted for clarity.)}
				\label{fig:hierarchy}
    		\end{figure}

        % ---------------------------------------------------------------------------- %
        %                                      IIT                                     %
        % ---------------------------------------------------------------------------- %
		\subsection{Integrated Information Theory}
            Integrated information theory (IIT) states that consciousness is integrated information and system's consciousness is determined by its causal properties \citep{tononi2016integrated}. \ac{OurTheory} is in line with IIT in that informational properties are thought to underlie consciousness. In this section, we will discuss \ac{OurTheory} in the light of IIT.

            \textbf{The concept of "information"}: In IIT, information refers to "integrated information":\\ \textquote[\citealt{tononi2016integrated}]{Information that is specified by a system that is irreducible to that specified by its parts.} In \ac{OurTheory}, information refers to "self-information", i.e. information about the states of conscious experience and the physical states of a process. Therefore, IIT focuses more on the relationships between consciousness and causal interactions among elements within a system, whereas \ac{OurTheory} focuses more on the informational relationships between conscious experience and being in a certain state of a process.

		    \textbf{The "Exclusion" axiom in IIT}: In IIT, the Exclusion axiom claims that of all overlapping sets of elements, only one set with maximal integrated information can be conscious. The exclusion axiom should be applied over elements, space, time, and scales \citep{oizumi2014phenomenology, hoel2016can}. \linelabel{line:multi-consciousness}Different from IIT, \ac{OurTheory} allows multiple consciousnesses coexist across different scales of coarse-graining within a system if they are informationally closed from each other. The two distinctive predictions decisively pinpoint the core concepts of the two theories.

		    \textbf{The concept of "integration"}: In IIT, integrated information is one of the core concept of defining conscious individuals. In the current paper, we did not include the notion of integrated information within \ac{OurTheory}. This, however, results in one of the current weaknesses of \ac{OurTheory} that it lacks the ability to individuate NTIC processes in some extreme cases (i.e., the problem of individuality). We discussed this weaknesses in Sec.~\ref{sec:Limitation and Future work}.

		    \textbf{Prediction after system damaged}:\linelabel{line:r2-cutting} \ac{OurTheory} and IIT lead to different predictions when a system suffers from a damage. For example, considering a densely connected networks whose dynamics forms an NTIC process. If we cut the network in half, IIT would predict that this results in two consciousnesses because elements in both networks still maintain high degrees of interactions. In contrast, \ac{OurTheory} would predict that this operation may completely destroy informational closure of the network and rendering both parts unconscious.

        % ---------------------------------------------------------------------------- %
        %                            Predictive Processing                             %
        % ---------------------------------------------------------------------------- %
		\subsection{Predictive Processing}
    		Predictive processing (PP) is a powerful framework which integrates several ideas from neuroscience. This emerging theoretical framework posits that neural systems constantly generate predictions about incoming sensory signals and updates predictions based on prediction errors between predictions and sensory signals. According to PP, neural systems constantly perform unconscious statistical inference about hidden causes in the external environment. The perceptual contents are the "best guess" about the environment states including these hidden causes \citep{clark_2013, Hohwy2013}. PP is well integrated with Bayesian brain hypothesis and has been used to interpret conscious perception in many domains \citep{Hohwy2013, seth2014predictive}.

    		%% The gaps between pp and consciousness
    		PP is a powerful explanatory framework for diverse brain functions. However, to serve as a theory of consciousness, PP is still incomplete due to two explanatory gaps. First, it has been known that the neural system is equipped multiple predictive mechanisms. Apparently, not all the predictive mechanisms are involved in conscious processes (e.g. mismatch negativity, \cite{naatanen2007mismatch}). PP needs to explain the difference between conscious and unconscious predictive mechanisms.

    		Second, PP can be considered as a sophisticated computation for perceptual inference. It takes von Helmholtz's conception of perception as unconscious inference. Thus, only the most probable outcome computed by the inference processes can be conscious while other details of the computation remain unconscious. PP also needs to explain how unconscious inferences is able to give rise to conscious results. In short, while PP is often discussed in the context of consciousness, these explanatory gaps prevent PP from being a theory of consciousness.

    		% compatible things
    		\ac{OurTheory} is well compatible with PP. Crucially, \ac{OurTheory} further provides natural and fundamental explanations to fills the two explanatory gaps which PP encounters. According to the definition of NTIC, a process with high NTIC can be regarded as a powerful predictive machine which has accurate self-predictive information ($I(Y_{t+1};Y_{t})$, E.q.~\ref{eq:NTIC1}) and concurrently incorporates environmental information into its dynamic ($I(Y_{t+1};Y_{t}|E_{t})$, E.q.~\ref{eq:NTIC1}). This predictive nature of NTIC processes is in agreement with the core notion of PP in which the conscious contents are always the predicted (inferred) outcome of our predictive mechanisms. Second, due to the informational closure to the environment, the encoded information about its environment in an NTIC process can be seemed as "the best guess" about the external environment in the context of Bayesian inference.

    		% fill the gaps
    		So, eventually, why are some predictive information conscious and some are not? \ac{OurTheory} predicts that only the predictions generated from mechanisms involving the NTIC process are conscious. Note that predictive processes are not necessary to involve NTIC processes. A predictive process can make prediction about the future state of its environment based on the current sensory inputs. In this case, the the process is not informationally closed and could not be conscious.

            According to \ac{OurTheory}, we further propose that we can only be aware of the predictions from predictive processes due to informational closure to computational details of microscopic predictive processes. The macroscopic NTIC process only acquires the coarse-grained summary statistics of the microscopic processes. In other words, we predict that the computation of the statistical inferences of PP is implemented at microscopic (cellular) scales in the neural system.

    		% finally
    		Finally, we consider PP as an potential empirical implementation of NTIC processes. To maintain accurate information about the environment encoded in an NTIC process, one can open an information channel between the process and the environment for minimal information flow to correct the divergence between them. This proposal is compatible with PP which suggests that PP systems updates (corrects) the current estimations by computing prediction errors between predicted and real sensory inputs.

        % ---------------------------------------------------------------------------- %
        %                           Sensorimotor contingency                           %
        % ---------------------------------------------------------------------------- %
		\subsection{Sensorimotor Contingency}
    		Sensorimotor contingency (SMC) theory of consciousness proposes that different types of SMCs give rise to different characteristics of conscious experience \citep{o2001sensorimotor}. The theory radically rejects the view that conscious content is associated with internal representations of a system. Rather, the quality of conscious experience depends on agents' mastery of SMCs. SMC emphasises that the interaction between a system and its environment determines conscious experience.

    	    \ac{OurTheory} is not compatible with SMC. As mentioned in Sec.~\ref{sec:Conscious versus Unconscious Processing}, a process directly maps the sensory states to the action states is insufficient to be NTIC. Therefore, learning contingencies between sensory inputs and action outputs does not imply NTIC. Hence, \ac{OurTheory} predicts that having sensorimotor contingencies is neither a necessary nor a sufficient condition for consciousness. In fact, empirically, with extensive training on a sensorimotor task with a fixed contingency, the task can be gradually performed unconsciously. This indicates that strong SMCs do not contribute conscious contents. In contrast, \ac{OurTheory} suggests that, with extensive training, the neural system establishes a neural mapping from sensory inputs to action outputs. This decrease the level of informational closure and, as a result, decrease the conscious level of this process. This outcome strongly supports \ac{OurTheory} than SMC.

    	    Nevertheless, \ac{OurTheory} does appreciate the notion that interactions between a process and its environment is crucial to shape conscious experience. As mentioned above, to form NTIC, a process needs to encode environmental transitions into its own dynamic. Therefore, information of agent-environment interaction should also be encoded in the NTIC process, and therefore, shape conscious contents in a specific way.

    	    Different from the classical SMC, a new version of SMC, Predictive Processing of SensoriMotor Contingencies (PPSMC), proposed by \cite{seth2014predictive, seth2015presence} combines SMC and the predictive processing framework together. PPSMC emphasises the important role of generative models in computing counterfactuals, inferring hidden causes of sensory signals, and linking fictive sensory signals to possible actions. According to \ac{OurTheory}, if the generative model involving the NTIC process for the computation of counterfactuals, PPSMC will be compatible to our theory and may have strong explanatory power on some specific conscious experience.

        % ---------------------------------------------------------------------------- %
        %                            Global workspace theory                           %
        % ---------------------------------------------------------------------------- %

		\subsection{Global Workspace Theory}
		Global workspace theory (GWT; \cite{baars1988cognitive, baars1997theatre, baars2002conscious}) or Global Neuronal Workspace theory (GNWT; \cite{dehaene1998neuronal, dehaene2001towards, dehaene2011experimental}) states that the neural system consists of several specialised modules and a central global workspace (GW) which integrates and broadcasts information gathered from those specialised modules. Only the information in the global workspace reaches conscious awareness, and information outside of it remains unconscious. These modules compete with each other to gain the access to the GW and the information from the winner triggers an all-or-none "ignition" in the GW. Information in the GW is broadcasted to other modules. Conscious contents then are associates with the information that gains access to the internal global workspace \cite{Dehaene2017}.

        While GWT emphasises the importance of global information sharing as a basis of consciousness, the precise meaning of information broadcasting has been somewhat unclear if one tries to describe it more formally in the language of information theory. \ac{OurTheory} offers one possible way to consider the meaning of broadcasting in GWT. Specifically, one could interpret the global workspace as the network of nodes where information is shared at the scale of NTIC where communication is performed through macro-variables that are linked via mutual predictability. That is, global workspace should be also NTIC. While this link remains speculative at this point, this interpretation encourages empirical studies into the relationship between the contents of consciousness and macrostate neural activities that are mutually predictive of each other.

% ============================================================================ %
%                                  Limitation                                  %
% ============================================================================ %
    \section{Limitation and Future Work}\label{sec:Limitation and Future work}  
        As a brand new theory of consciousness, \ac{OurTheory} is still far from completion. In the following, we discuss the current limitations and challenges of \ac{OurTheory} and point out the potential future research directions.

        % =============================== Theoretical  ===============================

        % Can't solve the hard problem
        It's important to clarify that \ac{OurTheory} does not intend to completely solve the hard problems of consciousness \citep{chalmers1995facing}. Knowing the state of a conscious process does not allow us to answer "What is it like to be in this state of this process" \citep{nagel1974like}. Instead, \ac{OurTheory} focuses more on bridging consciousness and the physical world using information theory as a common language in between.

        % Identity problem
        \linelabel{line:individuality}The current version of \ac{OurTheory} cannot entirely solve the problem of individuality in some extreme circumstances. In common cases, one can identify individual consciousnesses by computing the levels of NTIC of a process. This approach can also be applied to finding the boundaries of individual consciousnesses (for details of the boundary detection procedure, see \cite{krakauer2014information}). However, in some specific circumstances, individuality of consciousness is not clear. For instance, we can define a new process $Y$ and also its environment $E$ by recruiting two independent NTIC processes $Y^1~\&~Y^2$ and their environments $E^1~\&~E^2$, respectively. So that $Y = \{Y^1,Y^2\}$ and $E=\{E^1,E^2\}$. In such case, the new process $Y$ will also be NTIC to $E$. Therefore, the current version of \ac{OurTheory} cannot determine whether there are two smaller consciousnesses or one bigger consciousness (or 3 coexisting consciousnesses). The problem of individuality is a significant theoretical weakness of the current version of \ac{OurTheory}. The notion of integration is a possible remedy for this issue and we will address this issue more explicitly in our future work using the concept of synergy.

        % About the environment
        The current version of \ac{OurTheory} assume that consciousness is only contributed by non-trivial rather than trivial information encoded in a process. In other words, how much information about environmental states and dynamics encoded in a process is a key quantity for consciousness. However, we do not exclude the possibility that environmental information may be just a proxy of other informational quantities. More theoretical work is needed to elucidate the role of environments. This issue will also be discuss in our next theoretical paper.

        \linelabel{line:dream}Explaining conscious experience during dreaming is always a challenge to all the theories of consciousness. ICT currently does not have a certain answer to dreaming. However, we want to emphasise that not all the processes in the neural system are NTIC. This is trivial since some processes are evidently not informationally closed. They mainly passively react to sensory inputs or other processes in the neural system. To the conscious (NTIC) process, the rest of the neural system is part of the environment, and undoubtedly retains some degree of activity during sleep and dreaming. We speculate that, during dreaming, the neural system stably forms an NTIC process with respect to its environment, i.e. the other parts of the neural system. However, at the current stage, this is mere speculation. Searching for the NTIC process(es) during dreaming is a crucial step to extend the scope of ICT in future research.

        % ================================= Empirical ================================ 
        % finding the coarse-graining function 
        Empirically, a major challenge to \ac{OurTheory} is to find proper coarse-graining functions which map microscopic processes to macroscopic NTIC processes. This will become an imperative issue of finding neurological supporting evidence for \ac{OurTheory}. To find proper coarse-graining functions among infinite candidates \citep{price2007causation} seem to be very challenging. Nevertheless, there are still theoretical and technical progresses recently that may contribute to  solving this issue. For example, the concept of \textit{causal emergence} proposed by Hoel \citep{Hoel19790, Hoel2018} has been further developed recently. Causal emergence is highly relevant to the relationship between informational closure and coarse-graining. In their new study by \cite{klein2019uncertainty}, they started to compare how different coarse-graining functions influence causal emergence at macroscopic scales. \cite{PFANTE.2014, PFANTE.2014b} provides thorough mathematical analyses on level identification including informational closure. In neuroscience, the understanding of neural population codes also achieves a tremendous progress due to the advancement of recording technique and data science \citep{Kohn2016, panzeri2015neural}. \cite{Gamez2016} has also systematically described relevant issues in terms of finding data correlates of consciousness amount different levels of abstraction. We believe that interdisciplinary research is required to narrow down the scope of searching the coarse-graining functions and conscious processes at the coarse-grained scales in the neural system and beyond.

        \linelabel{line:r1-state-dep}In this article, we do not use a state-dependent formulation of NTIC. However, we believe that the state-dependent NTIC is essential to describe the dynamics of conscious experience. Therefore, further research using point-wise informational measures to construct state-dependent NTIC is needed for the next version of ICT.

        Finally another empirical challenge to \ac{OurTheory} is the  of empirical supporting evidence. This is understandable because the concept of NTIC is relatively new in the history of information science, not to mention in neuroscience. Very few experiments and data collections are designed for examining NTIC properties in neural systems. To our knowledge, only two studies \citep{Palmer2015, sederberg2018learning} coincidentally examine relevant properties in salamander retina. They found that the a large group of neural populations of retinal ganglion cells encoded predictive information about external stimuli also had high self-predictive information about their own future states. This result is in line with the characteristic of NTIC. We expect that there will be more empirical studies examining relevant neural properties of NTIC.

% ============================================================================ %
%                                  Conclusion                                  %
% ============================================================================ %
    \section{Conclusions}
    % short re-introduction 
    In this paper, we introduced \textbf{\acf{OurTheory}}, a new informational theory of consciousness. \ac{OurTheory} proposes that a process which forms \textbf{non-trivial informational closure (NTIC)} is conscious and through coarse-graining the neural system can form NTIC processes, i.e., conscious processes, at a certain macroscopic scales. \ac{OurTheory} considers that information is a common language to bridge the gap between conscious experience and the physical reality. Using information theory, \ac{OurTheory} proposes computational definitions for both conscious level and conscious content. This makes \ac{OurTheory} be able to generalise to any system beyond the human brains.

    % For neuroscience
    \ac{OurTheory} provides explanation for various findings from research of conscious and unconscious processing. The implications of \ac{OurTheory} point out that the scales of coarse-graining play a critical role in searching for neural substrates of consciousness. Improper measurements, e.g., too fine or too coarse in terms of the scale of measurements, of neurophysiological signals may lead to misleading results and misinterpretations.

    % For scientific theory of consciousness
    \ac{OurTheory} reconciles several theories of consciousness. \ac{OurTheory} indicates that they conditionally coincide with \ac{OurTheory}'s implications and predictions but, however, not the fundamental and sufficient conditions for consciousness. For example, theories includes the theories emphasising recurrent circuits \citep{lamme2006towards, edelman1992bright}, the theories highlighting the internal simulation,  predictive mechanism, and generative models ~\citep{revonsuo2006inner, clark_2013,Hohwy2013, kanai_chang_yu_de_abril_biehl_guttenberg_2019, seth2014predictive, seth2015presence}, and theories related to multilevel view of consciousness~\citep{pennartz2018consciousness,pennartz2015brain,prinz2007intermediate, jackendoff1987consciousness}. Notably, \ac{OurTheory} is proposed based on the non-functional hypothesis. Notwithstanding, its implications for the functional aspects of a system well fit several functionalist proposals.

	% For philosophy of mind
	Regarding philosophy of mind, \ac{OurTheory} connects several distinct arguments together. \ac{OurTheory} can be seen as an identity theory because it assumes a fundamental relation between consciousness and information. Second, the implications of \ac{OurTheory} tightly link consciousness to several cognitive functions in the context of evolution. This explains why people might intuitively have a functionalist point of view of consciousness. \ac{OurTheory} emphasises that informational closure between scales of coarse-graining is critical to form NTIC processes in some stochastic systems. In this case, especially for the neural system, forming conscious processes at the macroscopic scales coincide with the perspective of emergentism. Finally, forming NTIC (conscious) processes through many-to-one maps, i.e., coarse-graining, implies multiple realisability of consciousness. As a result, \ac{OurTheory} provides an integrated view for these arguments and is further capable of indicating how and why they are conditionally true.

	% Final 
	So far, the current version of \ac{OurTheory} is still far from completion. Further theoretical and empirical research is indispensably required for \ac{OurTheory} to improve and solve several issues in the current version. Nevertheless, \ac{OurTheory} offers explanation and prediction for consciousness science. We hope that \ac{OurTheory} provides a new way of thinking and understanding neural substrates of consciousness.

	% ============================================================================ %
	%                                      End                                     %
	% ============================================================================ %
	\section*{Acknowledgements}
	\addcontentsline{toc}{section}{Acknowledgements}
	A.C., Y.Y, and R.K. are funded by Japan Science and Technology Agency (JST) CREST project. Work by M.B. and R.K. on this publication was made possible through the support of a grant from Templeton World Charity Foundation, Inc. The opinions expressed in this publication are those of the authors and do not necessarily reflect the views of Templeton World Charity Foundation, Inc. This manuscript has been released as a Pre-Print at arXiv \citep{chang2019information}.

    \section*{Author Contributions Statement}
    \addcontentsline{toc}{section}{Author Contributions Statement}
    A.C. conceived and developed the theory. M.B. and A.C. contributed the mathematical formalisation of the theory. A.C., M.B, and R.K wrote the manuscript, based on a first draft by A.C. with extensive comments from Y.Y. All authors contributed to manuscript revision, read and approved the submitted version.

    \section*{Conflict of Interest Statement}
    \addcontentsline{toc}{section}{Conflict of Interest Statement}
    All authors were was employed by the company Araya Inc. The authors declare that the research was conducted in the absence of any commercial or financial relationships that could be construed as a potential conflict of interest.
\section{Appendix*}
\label{sec:appendix}
            Let us assume that the system only observes a part of the environment state.
            % \footnote{In theory this part of the environment state may still evolve deterministically over time in which case the system need only copy this part to achieve NTIC and consciousness. This seems to be a rare case as well and we assume the observed part of the environment is not close to deterministic in the following.} 
            % In this case it is well known that the observed part of the environment need not even be a Markov process even if the environment itself is deterministic. So the mutual information between subsequent observations may be much lower than the mutual information between subsequent environment states.

            % Furthermore, in this case the transfer entropy from environment to the system becomes equal to the transfer entropy from the observed part of the environment to the system.

            % Finally, note that the transfer entropy that can be achieved by copying the environment state or the observed part of the environment state equal to the mutual information between the subsequent environment states or observations. Copying the environment state corresponds to setting $S_{t+1}=E_t$ such that the transfer entropy is
            % \begin{align}
            %  I(S_{t+1}:E_t|S_t)=I(E_t:E_t|E_{t-1})=H(E_t|E_{t-1}).
            % \end{align}
            We can represent the part of the environment that we observe by the value of a function $f$ applied to the environment state. In this case we get for the transfer entropy
            \begin{align}
             I(S_{t+1}:E_t|S_t)=I(S_{t+1}:f(E_t)|S_t).
            \end{align}
            If the system only copies the observation we then get for the transfer entropy
            \begin{align}
             I(S_{t+1}:f(E_t)|S_t)=I(f(E_t):f(E_t)|f(E_{t-1}))=H(f(E_t)|f(E_{t-1}))
            \end{align}
            and for the mutual information
            \begin{align}
             I(S_{t+1};E_t)=I(f(E_{t+1});E_t)=H(f(E_t))
            \end{align}
            such that 
            \begin{align}
                NTIC_t(E\rightarrow S)=I(f(E_t);f(E_{t-1})).
            \end{align}
            This shows that whenever there is mutual information between subsequent observations a process that only copies the observations has positive NTIC. Note that any additional (internal) processing of the observation without reference to an additional internal state using a function $g$ can only reduce this mutual information:
            \begin{align}
               I(f(E_t);g(f(E_{t-1})))\leq I(f(E_t);f(E_{t-1})).
            \end{align}
            However, ignoring restrictions due to a possibly fixed choice of the universe process $X$ we find that for each such system there are other systems that achieve higher NTIC. For example define the system to be the "mirrored" and synchronized environment by setting $S_t:=E_t$. Then the transfer entropy vanishes
            \begin{align}
             I(S_{t+1}:E_t|S_t)=I(E_{t+1}:E_t|E_t)=0
            \end{align}
            and the mutual information is equal to the mutual information between current and next environment state:
            \begin{align}
                I(S_{t+1};E_t)=I(E_{t+1};E_t).
            \end{align}
             In the case where the environment has itself higher predictive mutual information than the observations it produces i.e.\ when
             \begin{align}                I(E_{t+1};E_t)\geq I(f(E_{t+1});f(E_t))
            \end{align}
            there is then potential for a predictive process to achieve higher NTIC than a copying system or any system that only processes its last observations without taking into account other internal memory (i.e.\ those systems also applying $g$ to their observations). Note that this also holds true in the case where the observations are themselves closed. If there is a more complex environment behind them, the mirrored and synchoronized system has higher NTIC with respect to that environment than the system copying the observations.

% ----------------------
%             For a feedforward network with each layer computing just the identity:
%             \begin{align}                NTIC&=I(f(E_t),\dots,f(E_{t-n+1});f(E_{t-1}),\dots,f(E_{t-n}))\\
%             &=H(f(E_t),\dots,f(E_{t-n+1})) - H(f(E_t)|f(E_{t-1}),\dots,f(E_{t-n})).
%             \end{align}
%             For the mirror environment, if we use the history dependent NTIC \citep{BERTSCHINGER.2006} we get:
%             \begin{align}
%                 NTIC^{s,m}=I(E_{t+1};E_t,\dots,E_{t-n+1})
%             \end{align}
%             To see that this is always larger or equal to the feedforward network note that the feedforward NTIC can be rewrittten:
%             \begin{align}                NTIC^{ff}&=I(f(E_t),\dots,f(E_{t-n+1});f(E_{t-1}),\dots,f(E_{t-n}))\\
%             &=I(f(E_t);f(E_{t-1}),\dots,f(E_{t-n}))+
%             I(f(E_{t-1}),\dots,f(E_{t-n+1});f(E_{t-1}),\dots,f(E_{t-n})|f(E_t))\\
%             &=I(f(E_t);f(E_{t-1}),\dots,f(E_{t-n}))+
%             I(f(E_{t-1}),\dots,f(E_{t-n+1});f(E_{t-1}),\dots,f(E_{t-n})|f(E_t))\\
%             &=I(f(E_t);g(E_{t-1},\dots,E_{t-n}))+
%             I(f(E_{t-1}),\dots,f(E_{t-n+1});f(E_{t-1}),\dots,f(E_{t-n})|f(E_t))\\
%             &=I(E_t;E_{t-1},\dots,E_{t-n}) \underbrace{-B+
%             I(f(E_{t-1}),\dots,f(E_{t-n+1});f(E_{t-1}),\dots,f(E_{t-n})|f(E_t))}_{\leq 0}\\
%             &\leq I(E_t;E_{t-1},\dots,E_{t-n}).
%             % &\leq NTIC^{s,m}+
%             % I(C;C,f(E_{t-n})|f(E_t))\\
%             % &=NTIC^{s,m}+
%             % I(C;C|f(E_t))\\
%             % &=NTIC^{s,m}+
%             % H(C|f(E_t))\\
%             % &=\underbrace{I(f(E_t);f(E_{t-1}),\dots,f(E_{t-n}))}_{\text{stationarity:}\leq NTIC^{s,m}}+
%             % I(f(E_{t-1}),\dots,f(E_{t-n+1});f(E_{t-1}),\dots,f(E_{t-n})|f(E_t))\\
%             % &\leq NTIC^{s,m}+
%             % I(C;C,f(E_{t-n})|f(E_t))\\            t
%             % &=NTIC^{s,m}+
%             % I(C;C|f(E_t))\\
%             % &=NTIC^{s,m}+
%             % H(C|f(E_t))\\
%             % ============================================================================ WEIRD
%             % &=I(f(E_t);C)+I(f(E_t),f(E_{t-n})|C)+
%             % H(C|f(E_t))\\
%             % &=H(C)-H(C|f(E_t))+I(f(E_t),f(E_{t-n})|C)+
%             % H(C|f(E_t))\\
%             % &=H(C)+I(f(E_t),f(E_{t-n})|C)\\
%             % &=H(C)+H(f(E_t)|C)-H(f(E_t)|f(E_{t-n}),C)\\
%             % &=I(f(E_t);C)-H(f(E_t)|f(E_{t-n}),C)\\
%             % &=I(A, C;C,B).
%             \end{align}

%             \begin{align}
%                 I(X;Y)&=_? I(f(X),X:g(Y),Y)\\
%                 &=I(X:g(Y),Y)+I(f(X):g(Y),Y|X)\\
%                 &=I(X:g(Y),Y)\\
%                 &=I(X:Y)+I(X:g(Y)|Y)\\
%                 &=I(X:Y)
%             \end{align}

%             \begin{align}
%                 I(X;Y)&= I(f(X),X:g(Y),Y)\\
%                 &=I(f(X):g(Y),Y)+I(X:g(Y),Y|f(X))\\
%                 &=I(f(X):g(Y))+I(f(X):Y|g(Y))+ I(X:Y|f(X))+I(X:g(Y)|Y,f(X)).
%             \end{align}

%             \begin{align}
%                 I(f(X):g(Y))=I(X;Y)-I(f(X):Y|g(Y))- I(X:Y|f(X))-I(X:g(Y)|Y,f(X))
%             \end{align}

% ----------------------
	\bibliographystyle{authordate1}
	\addcontentsline{toc}{section}{Reference}
	\bibliography{ref}

	\section*{Figure Legends}
	\renewcommand{\listfigurename}{~}
	\addcontentsline{toc}{section}{Figure Legends}
    \listoffigures

\end{document}